A pandemic-scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-09-26 latex/main.tex
Updates from Overleaf

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed a more comprehensive mutational landscape
of SARS-CoV-2 by mapping deletions onto the ARG
using post-hoc parsimony.
We identified important deletion events
(which are inherited by at least 10,000 samples;
Table~\ref{tab:major_dels}).
All of these deletions
(except a one-base deletion at 28,271
above a B.1-labeled node)
are lineage-defining mutations of major VOCs
(Alpha, Delta, Omicron BA.1, BA.2, and BA.4).
Many of these deletions occur in S1-NTD,
a known hotspot of in-frame deletions (REF).
There is a recurrent 1-base deletion at 28,271,
which is in a non-coding region.
<TODO: The 5-base dels and 4-base del are split up.
It should be a 9-base del (11,288–11,296).
I don't see excluded sites among these deleted sites.>

We also found highly recurrent deletions using the ARG.
The top 20 deletions in the number of occurrences
are shown in Table~\ref{tab:rec_dels}.
About half of these deletions affect a single base
and occur in non-coding regions
(concentrated in the 3' end of the genome);
some occur in accessory genes ORF7a and ORF8.
<TODO: Are any of these assembly errors or misalignments?>.
Two of these are known recurrent deletions in S1-NTD,
which are associated with the origins of major VOCs (REF).

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed a more comprehensive mutational landscape
of SARS-CoV-2 by mapping deletions onto the ARG
using post-hoc parsimony.
We identified important deletion events
(which are inherited by at least 10,000 samples;
Table~\ref{tab:major_dels}).
All these deletions are known to occur major VOCs
(https://virological.org/t/preliminary-genomic-characterisation-of-an-emergent-sars-cov-2-lineage-in-the-uk-defined-by-a-novel-set-of-spike-mutations/563;
\cite{Akaishi2023,Li2023indels}).
Eight of these deletions occur in S1-NTD,
a known hotspot for recurrent deletions \cite{McCarthy2021}.
""" The 5-base dels and 4-base del are split up as separate events.
It should be a 9-base del (11,288–11,296)."""
Furthermore, these deletion events reflect
the VOC-associated deletion events
inferred in a previous study \cite{Akaishi2023}.

We also found highly recurrent deletions using the ARG.
The top 20 deletions in the number of occurrences in the ARG
are shown in Table~\ref{tab:rec_dels}.
About half of these deletions are one-base,
occur in non-coding regions, and
are concentrated in the 3' end of the genome.
Four one-base deletions occur
in accessory genes (ORF7a, ORF8, and ORF10).
The two deletions in S1-NTD are known recurrent deletions
associated with antibody escape \cite{McCarthy2021}.
